---
title: "R Markdown Example of MS Word"
author: "Jason Thomas"
date: "01/08/2020"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)
# set default value for code chunks to print the code (along with the 
# output) in the MS Word document
opts_chunk$set(echo = TRUE)
```

# Markdown

Surround text with two asterisks, \*\*like this\*\*, to make
text **bold**.  Use a single asterisk, \*like this\*, to put
text in *italics*.  Here is a nested list (indentation is used
for sub-levels):

* Point A
  + A.1
    - A.1.1
	- A.1.2
	  * A.1.2.1
  + A.2
* Point B

You can include footnotes.^[To format footnotes (e.g., set the font size) you
probably have to use a template.  [Here](https://rmarkdown.rstudio.com/articles_docx.html) is 
an example.]  If you want use a `font that looks like computer code`^[I think it is called 
`Consolas` in MS Word.], surround your text with back ticks, i.e.,  \`like this\`.

## R Markdown

R Markdown is a special type of markdown that weaves R output directly into
the final document.  Here is an example.  

```{r cars}
summary(cars)
```

We can also include R output in a line.  For example, the third observation
of the speed variable is `r cars$speed[3]`.  Now we will replace it with a random
value

```{r cars2}
cars$speed[3] <- rnorm(1)
cars$speed[3]
```

and show that it has changed to `r cars$speed[3]`.

# Tables
```{r table}
# create table from iris data set (150 observations of plant species
# with variables sepal length and width, and pedal length and names.
names(iris)
table(iris$Species)

# create a matrix of means
tab1 <- aggregate(. ~ Species, data = iris, FUN = mean)
specieNames <- gsub("\\.", " ", names(iris))

# use kable() to format the table
kable(tab1, digits = 2, col.names = specieNames, align = c('l', rep('c', 4)),
      caption = "Table 1: Variable means for iris data.")
```

[This link](https://www.rdocumentation.org/packages/knitr/versions/1.26/topics/kable)
contains the help file for the `kable()` function.  



## Example with [kableExtra](https://cran.r-project.org/package=kableExtra)

Another useful R package is [kableExtra](https://cran.r-project.org/package=kableExtra).


## Example with [stargazer](https://cran.r-project.org/package=stargazer)

If you are interested in make tables for your regression models, take a
look at the R package [stargazer](https://cran.r-project.org/package=stargazer).


# Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

