---
title: "R Markdown Example of MS Word"
author: "Jason Thomas"
date: "01/14/2020"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)
library(summarytools)
library(furniture)
# set default value for code chunks to print the code (along with the output) 
# in the MS Word document
opts_chunk$set(comment = NA, echo = TRUE, results = "asis")
st_options(plain.ascii = FALSE, style = "rmarkdown",
           display.labels = FALSE, headings = FALSE)
regTab <- function(..., digits = 3) {

    modelsList <- list(...)
    rndCoef <- digits - 1
    # include check to ensure modelsList consists of "lm" objects

    nModels <- length(modelsList)
    nTerms <- lapply(modelsList, function(x) length(x$coef))
    modelOrder <- order(unlist(nTerms))
    modelMaxTerms <- which.max(nTerms)
    termsNames <- names(modelsList[[modelMaxTerms]]$coef)

    coefList <- lapply(modelsList,
                       function(x) format(coef(x), digits = rndCoef))

    coefList <- lapply(coefList, function(x) x[termsNames])
    coefMatrix <- matrix(unlist(coefList), ncol = nModels)
    coefMatrix <- coefMatrix[, modelOrder]
    coefMatrix[is.na(coefMatrix)] <- ""

    # add (formatted) row and column names
    termsNames <- gsub("\\(Intercept\\)", "intercept", termsNames)
    termsNames <- gsub("factor\\(([A-z]+)\\)([0-9]+)",
                       "\\1:\\2", termsNames)
    rownames(coefMatrix) <- termsNames
    colnames(coefMatrix) <- paste0("**Model ", 1:nModels, "**")

    return(coefMatrix)
}
```

# Introduction

This document shows how to use R Markdown to generate an MS Word document.
First, you will need to install pandoc, which can be downloaded at
[https://pandoc.org/installing.html](https://pandoc.org/installing.html).
Several R packages are also needed to create the final .docx file.  They
can be installed by running the following code in R Studio:

```{r install packages, eval = FALSE}
install.packages(c('rmarkdown', 'summarytools', 'furniture'))
```

# Markdown

Surround text with two asterisks, \*\*like this\*\*, to make
text **bold**.  Use a single asterisk, \*like this\*, to put
text in *italics*.  Here is a nested list (indentation is used
for sub-levels):

* Point A
  + A.1
    - A.1.1
	- A.1.2
	  * A.1.2.1
  + A.2
* Point B

    

You can include footnotes.^[To format footnotes (e.g., set the font size) you
probably have to use a template.  [Here](https://rmarkdown.rstudio.com/articles_docx.html) is 
an example.]  If you want use a `font that looks like computer code`^[I think it is called 
`Consolas` in MS Word.], surround your text with back ticks, i.e.,  \`like this\`.  Finally,
if you want to insert a new line, simply include `\` at the end of the line, as I do here \
and you can see that this second sentence fragment is on a new line \
and this third section is also on a new line.


    
## R Markdown
    

R Markdown is a special type of markdown that weaves R output directly into
the final document.^[More specifically, it actually call on another R package,
knitr, to incorporate the R output.]  Use code chunks to include R output, which are created
using three back ticks and some options surrounded by curly braces, like this:

\`\`\`{r cars}

summary(cars)
         
\`\`\`

This r code chunk (labeled cars) produces the following output in the MS Word document:

```{r cars}
summary(cars)
```

Note that each code chunk must have a unique name (or R will complain about 
'duplicate labels') and the document will fail to compile.  We can also 
include R output in a line.  For example, the third observation of the 
speed variable is `r cars$speed[3]`.  Now we will replace it with a 
random value

```{r cars2}
cars$speed[3] <- rnorm(1)
cars$speed[3]
```

and show that it has changed to `r cars$speed[3]`.


    
# Tables
    

There are several options.  The first is to use markdown syntax, but this is tedious and
only practical for very basic tables.  R packages, e.g.,
[summarytools](https://cran.r-project.org/package=summarytools) and 
[furniture](https://cran.r-project.org/package=furniture) are quite useful for making
tables, as illustrated below.

\

## Markdown syntax

\
\
\

| Col 1 | *Col 2* | **Column 3**|
|-------|:-------:|------------:|
| left justified | center |  right justified  |

\
\
\
\
or something like this
\
\
\
\
    
**Countries**   **Visited**     **Good Football team?**
-------------   -----------     ----------------------
G. Britain      Yes             Sometimes :)
India           Yes             No Asian Cup trophies yet
Zambia          Not yet         1 Cup of Nations Title 


\

    
But this is a pain!


\


## R packages


\
    

This section includes examples with [kable](https://www.rdocumentation.org/packages/knitr/versions/1.26/topics/kable), 
and [furniture](https://cran.r-project.org/package=furniture).  


\


There are some other options as well (but I think `kable()` and the furniture tools are best for MS Word)

* [summarytools](https://cran.r-project.org/package=summarytools) 
[(vignette)](https://cran.r-project.org/web/packages/summarytools/vignettes/Recommendations-rmarkdown.html)
* [pander](https://cran.r-project.org/package=pander)
[(vignette)](http://rapporter.github.io/pander/knitr.html)


\


### *Example with kable*
    

\


This examples uses the Iris dataset that is included with the base installation of R.


\

```{r table1}
# create table from iris data set (150 observations of plant species
# with variables sepal length and width, and pedal length and names.
names(iris)
table(iris$Species)

# create a matrix of means
tab1 <- aggregate(. ~ Species, data = iris, FUN = mean)
specieNames <- gsub("\\.", " ", names(iris))

# use kable() to format the table
kable(tab1, digits = 2, col.names = specieNames, align = c('l', rep('c', 4)),
      caption = "Table 1: Variable means for iris data.")
```


\

[This link](https://www.rdocumentation.org/packages/knitr/versions/1.26/topics/kable)
contains the help file for the `kable()` function.  

    
\


### *Example with furniture*


\


First, we'll create a table of descriptive statistics using
`table1()`


\


```{r descriptives furniture ex1}
table1(mtcars, type = c("simple", "condensed"),
       output = "pandoc")
```


\


Another example, but for a subset of variables and different
summary functions.


\


```{r descriptives furniture ex2}
vNames <- c("Miles per gallon", "Number of cylinders",
            "Gross horsepower")
table1(mtcars, mpg, cyl, hp, var_names = vNames,
       caption = "**Means and SD for subset of variables**",
       type = c("simple", "condensed"),
       output = "pandoc")
```


\


```{r descriptives furniture ex3}
vNames <- c("Miles per gallon", "Number of cylinders",
            "Gross horsepower")
table1(mtcars, mpg, cyl, hp, var_names = vNames,
       splitby = ~am,
       caption = "**Means and SD for subset of variables**",
       type = c("simple", "condensed"),
       output = "pandoc")
```


\


```{r cross tab tableX}
crossTab <- tableX(mtcars, cyl, am, type = "count")
kable(crossTab)
```


\


### *Example with regression tables*
    

```{r regression table}
model1 <- lm(mpg ~ hp, data = mtcars)
model2 <- lm(mpg ~ hp + drat, data = mtcars)
model3 <- lm(mpg ~ hp + drat + factor(gear), data = mtcars)
tab2 <- regTab(model1, model2, model3)
kable(tab2, align = c('l', rep('c', 3)))
```

    
    


    
# Including Plots
    

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

